---
title: "Confusion"
author: "Lyron Winderbaum"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Contents

- Statistical Tests.
- Confusion Matrix.
- Family-wise Error Rate.
- R Markdown.

## Statistical Tests

What is a statistical test? 

In this context, a test is a method for answering a 
binary question --- a question with two possible outcomes.

We will use the terminology Positive/ Negative for these two outcomes although 
different terminology might be appropriate in any given context.

Some examples:

- A test for HIV can result in positive/ negative results.
- Similarly a test for LNM.
- Testing for the difference in the mean serum concentration of a protein 
  between two experimental groups results in either a difference being detected, 
  or no difference detected.

## Confusion Matrix

-------------------- -------------------- 
 TP: True  Positive   FN: False Negative
 FP: False Positive   TN: True  Negative 
-------------------- -------------------- 

This (possibly) deceptive 2x2 table is called the confusion matrix.

The words "True" and "False" refer to the truth, while the words "Positive"
and "Negative" refer to the outcome of your test.

## Confusion Matrix
Many common concepts are directly derived from the confusion matrix.
Including many TLA (Three Letter Acronyms) that you should recognise.

- Significance: FPR = FP / (FP + TN) = 1 - specificity
- Power: TPR = TP / (TP + FN) = sensitivity
- Specificity: TNR = TN / (FP + TN))
- Prevalence: (TP + FN) / (TP + FN + FP + TN)
- Precision: PPV = TP / (TP + FP)
- False Discovery Rate: FDR = FP / (TP + FP) = 1 - PPV

Notice Significance and Power are plotted against each other in an ROC curve.

## Confusion Matrix
And some that you might not recognise.

- Accuracy: (TP + TN) / (TP + FN + FP + TN)
- F1 = 2TP / (2TP + FP + FN)
- Informedness: sensitivity + specificity - 1 = DIPPS

and in case you're curious and interested in further reading,
you might like to check out the 
[Mathews Correlation Coefficient](https://en.wikipedia.org/wiki/Matthews_correlation_coefficient) (MCC). 

If you do follow up by reading more about this topic, note
that my notation here is cosistent with that of the 
[wikipedia page](https://en.wikipedia.org/wiki/Sensitivity_and_specificity).

## P-Values
But for now lets just focus on one thing: Significance, also known as False 
Positive Rate (FPR), or type I error rate.

Why?

Because we were talking about statistical tests, and the first thing that brings
to mind is p-values. 

What is the link between p-values and FPR?

Remember where we started? A statistical test is a method for producing an 
answer to a binary question. Yes, or no. Often, this is acheived by calculating 
a p-value, and then choosing an answer based on if the resulting p-value is less
than 0.05 or not.

Why do we make decisions in this way?

## P-Values

Well, because assuming that you have calculated your p-values correctly, and 
that any required assumptions are true, then making our decision in this way 
will produce a FPR of 0.05.

What does that mean?

Well, lets take an example. Lets say you are looking for a difference in the 
concentration of a protein between two groups. In that context, it means that if 
you where to replicate your experiement many, many times, you would 
expect 5% of the replicates in which the truth was that there was no difference 
to produce the opposite conclusion (that there was infact a difference).

## Family-wise Error Rate

Now that was under the assumption you were just looking at a single protein. 
What if you were looking at several proteins? Well, the FPR, in this context
interpreted as the probability of coming to the conclusion that there is a 
difference in protein concentration when in truth there is no such difference, 
would still be 5% for each individual protein independantly, again assumping 
p-values where calculated correctly and all neccessary assumptions are 
satisfied.

However, the probability of there being at least one false positive amongst 
the set of tests performed (each with individual probability 5%), will be much 
greater than 5%. This is called the Family-wise error rate (FWER). Often 
attention will be focused on differences found, which will naturally 
include such False Positives that occur with increased probability, and this can 
be problematic.

## Multiple Testing Correction
It can however be addressed by adjusting the p-value cutoff used in the 
decision-making process, thereby controlling the FWER. 

In practice, this results in using more stringent p-value cutoffs in order to 
guarantee a 5% FWER. 

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for
authoring HTML, PDF, and MS Word documents. For more details on using R Markdown
see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes
both content as well as the output of any embedded R code chunks within the
document. A couple of examples follow.

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```
